---
title: "Siplr_Usages"
author: "Emmanuel BATT"
date: "26/07/2021"
output: html_document
---

```{r setup, include=FALSE}
library(swiplr)
swipl <- swiplr::swiplR()

```


## Tests databases

### Test 1 for upper case and spécial caracters in atoms

```{prolog test1_data}

test1(foo, bar).
test1(foo, 'Bar').
test1(foo, 'Bar of joe.').
test1('_foo_', 'Bar of Joé.').
test1(fée, 'Bar of Joé.').

test1(with, more, columns, in, order, 'to hide', some, of, them).
test1(with, less, columns, in, order, 'to show', many, of, them).

```


### Test 2 complex data in results

```{prolog test2_data}

test2(some(foo_bar),other(with,things)).
test2(some(of(lists,number)),[1,2,3]).
test2(some(of(lists,things)),[some(thing),2,foo(bar)]).

```


## Query test databases

### Test 1 for special characters

```{r}
swipl$query(.swiplr_chunks$test1_data, 
                    query = "test1(Foo, Bar)")
```

Warning: #€$ and () are not allowed in Variable names

```{r}
swipl$query(.swiplr_chunks$test1_data, 
                    query = "test1(FoéWithÀccent, Bar_With_Ümlletter)")
```

```{r}
swipl$query(.swiplr_chunks$test1_data, 
                    query = "test1(FoéWithÀccent, Bar_With_Ümlletter)")
```

Hiding columns after query 

```{r}
swipl$query(.swiplr_chunks$test1_data, 
                    query = "test1(Visible, Hidden_)")
```

```{r}
swipl$query(.swiplr_chunks$test1_data, 
                    query = "test1(Hidden_, Visible, A, B, C, D, E, F, G)")

swipl$query(.swiplr_chunks$test1_data, 
                    query = "test1(Hidden_, Visible, A_, B_, C_, D, E_, F_, G_)")

swipl$query(.swiplr_chunks$test1_data, 
                    query = "test1(Hidden_, Visible, _, _, _, D_, _, _, _)")

```


### Test 2 : Multi level query

Automatic type unlist for columns

```{r}
swipl$query(.swiplr_chunks$test2_data, 
                    query = "test2(What, Content)")


swipl$query(.swiplr_chunks$test2_data, 
                    query = "test2(some(foo_bar), Content)")

```

```{r}

l_out <- swipl$query(.swiplr_chunks$test2_data, 
                    query = "test2(some(of(lists, What)), List)")

str(l_out$List)
```

### Test 3 : Inject data in body

```{r}
body <- "
test3(some({{{key}}}_always_visible),other({{{value}}})).

{{{#visible}}}
test3(some({{{key}}}_may_hide),other({{{value}}})).
{{{/visible}}}
"


swipl$query(body, query = "test3(Show, Other)",
  data = list(key="foo1", value="bar1é", visible=T))

swipl$query(body, query = "test3(Hide, Other)",
  data = list(key="foo2", value="bar2é", visible=F))

```


```{r}
body <- "
{{{#row}}}
test3(some({{{key}}}),other({{{value}}})).
{{{/row}}}
"
  
swipl$query(body,
  query =   "test3(Key, Value)",
  data = list(row=list(
              list(key="foo1", value="bar1"),
              list(key="foo2", value="bar2")
              ) )
  )

```

This is how to convert a data.frame into a prolog predicate.

```{r}
body <- "
{{{#row}}}
co2('{{{Plant}}}', '{{{Type}}}', {{{conc}}}, {{{uptake}}}).
{{{/row}}}
"
data("CO2")
swipl$query(body,
  query =   "co2(Plant, Type, Conc, Uptake)",
  data = list(row=unname(apply(CO2[1:5,], 1, as.list) ) )
  )

```

## Broken files and error handling

### Test 4: bad syntax in file


```{r}
swipl$restart()

swipl$query("missing_point(at,end,of,line)", "X=1")
swipl$query("missing_operator ponctuation.", "X=1")
swipl$query(":- my_undefined_procedure.", "X=1")

swipl$query("foo(bar).", "foo(X)")

```

### Test 5: bad syntax in query


```{r}
swipl$restart()

swipl$query("foo(bar).", "foo(X ")
swipl$query("foo(bar).", "foo2(x) ")

```


## Performances

### full call duration

```{r}
# usually  less than 100ms
system.time({out <- swiplr::swiplR()$query("foo(bar).", "foo(X).")})
out

pl <- swiplr::swiplR()
# usually  less than 10 ms
system.time({out <- pl$query("foo(bar).", "foo(X).")})
out


```


```{r}
data <- as.pl_data.data.frame(CO2)
pl <- swiplr::swiplR()
# usually  less than 80 ms
system.time({out <- pl$query(
  "\n{{{#co2}}}co2('{{{Plant}}}','{{{Type}}}','{{{Treatment}}}','{{{conc}}}',{{{uptake}}}).\n{{{/co2}}}",
  "co2(Plant, Type, Treatment, Conc, Uptake).", 
  data = data, nsol = 1000)})

out

data <- as.pl_data.data.frame(EuStockMarkets)
# usually 1500 ms
system.time({out <- pl$query(
  "\n{{{#eustockmarkets}}}eustockmarkets('{{{DAX}}}','{{{SMI}}}','{{{CAC}}}','{{{FTSE}}}').\n{{{/eustockmarkets}}}",
  "eustockmarkets(DAX, SMI, CAC, FTSE).", 
  data = data, nsol = 2000)})
out
```
